---
title: "Analytic Plan Feb 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Overall Considerations

In making sure my analyses for ACToday are optimized, I wanted to ask:

  * What analyses are feasible within my limited time remaining with ACToday?
  * What information would be most useful for the ACToday project, given its focus on climate services and policy change?

### Research Questions

  1. Do extreme climate events increase prevalence of wasting among children in Bangladesh?
    a. If so, are there lags that are important to these relationships?

### Methodological Considerations

#### Case-Crossover Design

Case-crossover studies usually require (quoted from [Columbia Mailman SPH](https://www.mailman.columbia.edu/research/population-health-methods/case-crossover-design)):

  * Short-term changes in exposure
  * Transient changes in disease risk
  * Acute-onset disease
  
My main questions with regards to our data:

  * What is the time threshold at which disease risk becomes 'transient'? Can we consider wasting, with a lag period of weeks to potentially months, transient?
  * Similarly, there is uncertainty surrounding the actual 'onset' of wasting. Would it make sense to investigate these extreme events as 'triggers' that tip a child to acute malnutrition?
  * Finally, in a case-crossover analysis of these data, would these potential lag structures interfere with our ability to select our case and control windows?
  
#### Log-Binomal Analysis of Cross-Sectional Data

Similar to our previous analyses, these would entail comparison of wasting prevalences across differing levels of exposure, adjusting for confounders. This would produce adjusted Prevalence Ratios.

Questions here:

  * How do we investigate lags here - should it be a formal hypothesis testing process, or included in our exploratory analyses?
  * We've been looking at extreme precipitation and temperature events, but less so at anomalously low values. Do we want to include these as well (e.g. drought)?

#### Other Methods

Have also considered:

  * Distributed lag nonlinear models (DLNM) 
    + What would this entail given our data structure?
  * Time-series models (likely Poisson or quasi-Poisson regression?)
    + Not necessarily mapping acute exposure/outcome but overall
    + Could incorporate lags here
    + Could also consider season-stratified analyses
    
### Exploratory Analyses

I tried exploring the potential lag structures of temperature and heat anomalies. 

Adjusted log-binomial models were run as follows (basic example code):

```{r, eval = FALSE}
logbin(
  data = data,
  formula = y ~ x +    ## substitute vector of lags for x
    surv_area + moy + year
)
```

Graphs of results, adjusted by year, month, and geographic region, are below.

```{r, include = FALSE, echo = FALSE}
library(tidyverse)
library(logbin)

## saveRDS(test_results_adj, file = "./logbin_results/test_adj.RDS")
test_results_adj2_t =
  readRDS(file = "./logbin_results/test_adj.RDS")

## saveRDS(test_results_adj_p_pos, file = "./logbin_results/test_adj_p_pos.RDS")
test_results_adj_p_pos =
  readRDS(file = "./logbin_results/test_adj_p_pos.RDS")

## saveRDS(test_results_adj_p_neg, file = "./logbin_results/test_adj_p_neg.RDS")
test_results_adj_p_neg =
  readRDS(file = "./logbin_results/test_adj_p_neg.RDS")
```

```{r plots, fig.width = 12}
test_results_adj2_t %>% 
  mutate(
    beta1_df = as.ordered(beta1_df),
    time = 0:16,
    beta1_df = fct_reorder(beta1_df, time)) %>% 
  ggplot(
    aes(x = beta1_df,
        y = exp(estimate),
        ymin = exp(lower),
        ymax = exp(upper))
  ) +
    geom_point() +
    geom_errorbar() +
    geom_hline(yintercept = 1) +
    theme_light() +
    labs(title = "Weekly *Positive* Temperature Anomaly (C)\nAnomaly calculated using observed mean weekly value minus the long-term mean over time period for that week of year",
         x = "Lag",
         y = "Prevalence Ratio + 95% CI") 

test_results_adj_p_pos %>% 
  mutate(
    beta1_df = as.ordered(beta1_df),
    time = 0:16,
    beta1_df = fct_reorder(beta1_df, time)) %>% 
  ggplot(
    aes(x = beta1_df,
        y = exp(estimate),
        ymin = exp(lower),
        ymax = exp(upper))
  ) +
    geom_point() +
    geom_errorbar() +
    geom_hline(yintercept = 1) +
    theme_light() +
    labs(title = "Weekly *Positive* Precipitation Anomaly (mm)\nAnomaly calculated using observed mean weekly value minus the long-term mean over time period for that week of year",
         x = "Lag",
         y = "Prevalence Ratio + 95% CI") 

test_results_adj_p_neg %>% 
  mutate(
    beta1_df = as.ordered(beta1_df),
    time = 0:16,
    beta1_df = fct_reorder(beta1_df, time)) %>% 
  ggplot(
    aes(x = beta1_df,
        y = exp(estimate),
        ymin = exp(lower),
        ymax = exp(upper))
  ) +
    geom_point() +
    geom_errorbar() +
    geom_hline(yintercept = 1) +
    theme_light() +
    labs(title = "Weekly *Negative* Precipitation Anomaly (mm)\nAnomaly calculated using observed mean weekly value minus the long-term mean over time period for that week of year",
         x = "Lag",
         y = "Prevalence Ratio + 95% CI") 

```

Notes 02/12/2020

  * Continue with **log-binomial** models
  * What are plausible biological/other pathways that could lead to these lags?
    + High temperature
    + High precipitation
    + Low precipitation
  * What is onset period of wasting, particularly with non-climate ex. in literature (e.g. onset of wasting after famine)?
  * Operationalization of heat
  * EM by 
    + Season
    + Region
      * Politically actionable regions above survey area? --> district, level above district
  * Self-notes
    + Spline functions for seasonal/long-term patterns
    + Interaction term / random effect? for region
      
*Next Steps*

  * Literature searches:
    + Plausible pathways from climate exposures to wasting in children
    + Onset period of wasting from various insults (e.g. famine, non-climate but related)
  * Consider exposure operationalizations
  * Explore data:
    + Differential effects by region
    + Differential effects by season
    
# Notes

Pathways

  * Heat --> Diarrhea --> Wasting (Hashizume et al 2007, )
  * Heat --> Cholera --> Diarrhea/other --> Wasting (Checkley et al 2000, Lama et al 2004)
    + Lots on this - search for heat and diarrhea/cholera on Google Scholar
  
  * Precipitation --> Diarrhea --> Wasting ()
  * Precipitation --> Cholera --> Diarrhea --> Wasting (Hashizume et al 2008)
  
  * Other ideas
    + Health effects of ambient temperature (Xu et al 2012), diurnal temperature range (Cheng et al 2014), heat waves (Xu et al 2014) - all systematic reviews
    + Health effects of precipitation anomalies - search Google Scholar
    + Crop yields? (wouldn't see short-term fluctuations due to temp but maybe precip if at appropriate lag)
    + Food prices? (not sure about temp here, but precip maybe, similar to previous bullet)
    + Heat --> dehydration --> ??? --> wasting?
  
Timing

  * Rapid onset of wasting - no specific timeframe given (WFP 2005)
  * Rapid recovery - within 3 months (Rivera and Habicht 1996)
  
